---
title: "Master notebook for metaRL project"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---
The settings for the script are below, note that ntrials is reset if the task is loaded in. Save after running to create a new notebook each time. 
```{r}
#clears environment
rm(list=ls())

#testing is TRUE means runs with less of everything for speed
testing=FALSE

# settings for script run 
newtask=0 #set to 1 to create a new task
taskname='task1' #either creates a task with this name, or loads in one with this name
ntrials=100 #trials in task
probability=0.75 #probability of each shape being rewarded/punished
newchoices=1 #set to 1 to generate new participant choices
participant_multiplier=10 #set to 10 to have ten times the participants in the original study, etc
if (testing==TRUE) participant_multiplier=1

# directories
workingdir='N:/Alex/metaRL'
scriptdir='N:/Alex/metaRL/scripts'
datadir='N:/Alex/metaRL/simulated_data'
taskdir='N:/Alex/metaRL/task/'

# stan parameters
#stanname='metaRL_2a1b.stan'
chain_iter=2e4 # number of iterations for each chain inc. warmup (half)
if (testing==TRUE) chain_iter=2e3
chain_n=4 #4 chains
if (testing==TRUE) chain_n=1
```
This loads the libraries and source files needed to run this script, and sets up RStan
```{r}
# libraries and source files 
library('MASS')
library('boot')
library('matlab')
library('dplyr')
library('reshape')
library('tidyr')
library('rstan') 
library('loo')

#options for RSTAN
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
Sys.setenv(LOCAL_CPPFLAGS = '-march=native')
Sys.getenv('LOCAL_CPPFLAGS') #should say '-march=native'

#functions
source('N:/Alex/metaRL/scripts/simulate_RL.R') #creates a normal distribution around params
source('N:/Alex/metaRL/scripts/simulate_choices.R') #simulates participant choices using Q learning and softmax
source('N:/Alex/metaRL/scripts/create_task.R') #creates task if necessary
source('N:/Alex/metaRL/scripts/estimate_params.R') #recovers parameters
source('N:/Alex/metaRL/scripts/create_sim_matrix.R')#creates the matrix of simulation outputs
```
This section of code creates a task (if the initial settings are newtask==1), or loads in the task values from previous runs. Typically, the task has 10,000 trials and 1 reversal, with a probability of 0.8. This is generated using the script 'create_task', which creates a simple reversal learning task. Wins and losses are coupled - if you don't 'win' by choosing one option, you lose; and if the shape you chose doesn't win, the other does. 

```{r}
# creates new task, or loads previously saved one
ifelse(newtask==1, task<- create_task (ntrials, nreversals = 2, probability = probability), load(file.path(taskdir,taskname))) 
if(newtask==0) {ntrials <- nrow(task)} #if loads in a task, sets ntrials to be the number of trials in that task
save(task,file=file.path(taskdir,taskname))
```

Now, we simulate parameters from the Chase et al. 2009 paper. <br>
Glossary: <br> aG = positive learning rate <br> aL = negative learning rate <br>
tr = training phase <br> te = test phase <br> (not using test phase as this is PIT)
pat = patient <br> con = control

We use 10 times the number of participants in the original paper.
```{r}
# this part of the script simulates data from the Chase et al. 2009 paper

npat_Chase=23*participant_multiplier #number of participants in Chase paper * multiplier (defined at top)
ncon_Chase=23*participant_multiplier #ditto
nsub_Chase=npat_Chase+ncon_Chase #total subs

#probably going to ignore the test phase as not learning but PIT
#positive learning rate for patients in training phase
Chase_awin_pat <- simulate_RL (n = npat_Chase, est_m = 0.35, est_sd = 0.09) 

#positive learning rate for controls in training phase
Chase_awin_con <- simulate_RL (n = ncon_Chase, est_m = 0.23, est_sd = 0.06)

#negative learning rate for patients in training phase
Chase_aloss_pat <- simulate_RL (n = npat_Chase, est_m = 0.21, est_sd = 0.09)

#negative learning rate for controls in training phase
Chase_aloss_con <- simulate_RL (n = ncon_Chase, est_m = 0.24, est_sd = 0.08)

#exploration/exploitation (essentially beta) for patients in training phase
Chase_beta_pat <- simulate_RL (n = npat_Chase, est_m = 0.14, est_sd = 0.03)

#exploration/exploitation (essentially beta) for controls in training phase
Chase_beta_con <- simulate_RL (n = ncon_Chase, est_m = 0.13, est_sd = 0.02)

```
And generate data from the Kumar 2018 paper. 

```{r}
npat_Kumar=25*participant_multiplier #number of participants in Kumar paper * multiplier (defined at top)
ncon_Kumar=26*participant_multiplier #ditto
nsub_Kumar=npat_Kumar+ncon_Kumar #total subs for paper

#positive learning rate for patients 
Kumar_awin_pat <- simulate_RL (n = npat_Kumar, est_m = 0.28, est_sd = 0.14) 

#positive learning rate for controls 
Kumar_awin_con <- simulate_RL (n = ncon_Kumar, est_m = 0.31, est_sd = 0.14)

#negative learning rate for patients
Kumar_aloss_pat <- simulate_RL (n = npat_Kumar, est_m = 0.42, est_sd = 0.24)

#negative learning rate for controls
Kumar_aloss_con <- simulate_RL (n = ncon_Kumar, est_m = 0.38, est_sd = 0.23)

#positive beta for patients 
Kumar_betawin_pat <- simulate_RL (n = npat_Kumar, est_m = 0.13, est_sd = 0.15)

#positive beta for controls 
Kumar_betawin_con <- simulate_RL (n = ncon_Kumar, est_m = 0.11, est_sd = 0.16)

#negative beta for patients
Kumar_betaloss_pat <- simulate_RL (n = npat_Kumar, est_m = 0.19, est_sd = 0.20)

#negative beta for controls
Kumar_betaloss_con <- simulate_RL (n = ncon_Kumar, est_m = 0.23, est_sd = 0.17)
```

The below simulates patient choices using the simulated params from Chase et al. 2009
```{r}
#simulate 'participant' choices 

if (newchoices==1){
  temp1<- matrix(data=NA, nrow <- ntrials*(nsub_Chase), ncol<- 7)
  temp2<- matrix(data=NA, nrow <- ntrials*(nsub_Kumar), ncol<- 7)
  #patients Chase paper
  for (pat in 1:npat_Chase){
    choices<- simulate_choices (Chase_awin_pat[pat], Chase_aloss_pat[pat], Chase_beta_pat[pat],'empty', task) #null:no separate           betaloss
    temp1<-create_sim_matrix(temp1,1,1,pat,task,choices,ntrials)
    rm(choices)
  }
  #controls Chase paper
  for (con in 1:ncon_Chase){
    choices<- simulate_choices (Chase_awin_con[con], Chase_aloss_con[con], Chase_beta_con[con],'empty', task) 
    temp1<-create_sim_matrix(temp1,1,0,npat_Chase+con,task,choices,ntrials)
    rm(choices)
  }
  #patients Kumar paper
    for (pat in 1:npat_Kumar){
    choices<- simulate_choices (Kumar_awin_pat[pat], Kumar_aloss_pat[pat], Kumar_betawin_pat[pat],
                                Kumar_betaloss_pat[pat],task)
    temp2<-create_sim_matrix(temp2,2,1,pat,task,choices,ntrials)
    rm(choices)
  }
  #controls Kumar paper
  for (con in 1:ncon_Kumar){
    choices<- simulate_choices (Kumar_awin_con[con], Kumar_aloss_con[con], Kumar_betawin_con[con],
                                Kumar_betaloss_con[con],task)
    temp2<-create_sim_matrix(temp2,2,0,npat_Kumar+con,task,choices,ntrials)
    rm(choices)
  }
  simulated_data<-rbind(temp1,temp2)
  rm(temp1,temp2)
} else {
  load('N:/Alex/metaRL/simulated_data/simulated_data')
}

```
Now, we save the simulated data.
```{r}
save(simulated_data,file='N:/Alex/metaRL/simulated_data/simulated_data')

```
We want to see if the parameters are recoverable. Therefore, we need to go from these choices back to the parameters that generated them (this does this for the Chase paper).

```{r}
stanname='metaRL_Chase2.stan'
stanfile <- file.path(scriptdir, stanname)

simulated_data<-as.data.frame(simulated_data)
colnames(simulated_data)<- c('study','pat/con','id','trial','reward','pun','choices')
rm(rewardA,punA,choices)

simulated_data_Chase<-subset(simulated_data,simulated_data$study==1) #Chase data

rewardA <- select(simulated_data_Chase,c('id','trial','reward')) %>% spread(id, reward, fill = 0)
rewardA$trial <- NULL
punishA <- select(simulated_data_Chase,c('id','trial','pun')) %>% spread(id, pun, fill = 0)
punishA$trial <- NULL
choices <- select(simulated_data_Chase,c('id','trial','choices')) %>% spread(id, choices, fill = 0)
choices$trial <- NULL

Chase_data<-list(ntrials=ntrials,nsub=nsub_Chase,includeTrial = rep(1,ntrials),  rewardA=rewardA,punishA=punishA,choices=choices)

Chase_fit <- stan(file = stanfile, data = Chase_data, iter=chain_iter, chains = chain_n) #add working dir?
save(Chase_fit, file=file.path(datadir,'Chase_fit'))

traceplot(Chase_fit,'lp__')

# draws_Chase<- extract(Chase_fit)
summary_Chase<- summary(Chase_fit)

Chase_loglike<- extract_log_lik(Chase_fit, parameter_name = "loglik", merge_chains = TRUE)

loo(Chase_loglike)

overall_summary<-summary_Chase$summary #summary of all chains merged

#puts output in form for correlations
estimated_Chase_output<- matrix(data=NA, nrow <- nsub_Chase, ncol<- 6)
rownames<-paste('alpha[',1:nsub_Chase,',1]',sep='')
estimated_Chase_output[,1] <- overall_summary[c(rownames),'mean']
rownames<-paste('alpha[',1:nsub_Chase,',1]',sep='')
estimated_Chase_output[,2] <- overall_summary[c(rownames),'sd']
rownames<-paste('alpha[',1:nsub_Chase,',2]',sep='')
estimated_Chase_output[,3] <- overall_summary[c(rownames),'mean']
rownames<-paste('alpha[',1:nsub_Chase,',2]',sep='')
estimated_Chase_output[,4] <- overall_summary[c(rownames),'sd']
rownames<-paste('beta[',1:nsub_Chase,']',sep='')
estimated_Chase_output[,5] <- overall_summary[c(rownames),'mean']
rownames<-paste('beta[',1:nsub_Chase,']',sep='')
estimated_Chase_output[,6] <- overall_summary[c(rownames),'sd']

#for positive learning rate
cor.test(c(Chase_awin_pat,Chase_awin_con),estimated_Chase_output[,1]) #correlation
lm(c(Chase_awin_pat,Chase_awin_con)~estimated_Chase_output[,1]) #regression equation
plot(c(Chase_awin_pat,Chase_awin_con),estimated_Chase_output[,1]) #plot

#for negative learning rate
cor.test(c(Chase_aloss_pat,Chase_aloss_con),estimated_Chase_output[,3])
lm(c(Chase_aloss_pat,Chase_aloss_con)~estimated_Chase_output[,3])
plot(c(Chase_aloss_pat,Chase_aloss_con),estimated_Chase_output[,3])

#for beta
cor.test(c(Chase_beta_pat,Chase_beta_con),estimated_Chase_output[,5])
lm(c(Chase_beta_pat,Chase_beta_con)~estimated_Chase_output[,5])
plot(c(Chase_beta_pat,Chase_beta_con),estimated_Chase_output[,5])

```
And see if the parameters from Kumar paper are recoverable
```{r}
stanname='metaRL_Kumar.stan'
stanfile <- file.path(scriptdir, stanname)

simulated_data<-as.data.frame(simulated_data)
colnames(simulated_data)<- c('study','pat/con','id','trial','reward','pun','choices')
rm(rewardA,punishA,choices)

simulated_data_Kumar<-subset(simulated_data,simulated_data$study==2) #Kumar data

rewardA <- select(simulated_data_Kumar,c('id','trial','reward')) %>% spread(id, reward, fill = 0)
rewardA$trial <- NULL
punishA <- select(simulated_data_Kumar,c('id','trial','pun')) %>% spread(id, pun, fill = 0)
punishA$trial <- NULL
choices <- select(simulated_data_Kumar,c('id','trial','choices')) %>% spread(id, choices, fill = 0)
choices$trial <- NULL

Kumar_data<-list(ntrials=ntrials,nsub=nsub_Kumar,includeTrial = rep(1,ntrials),  rewardA=rewardA,punishA=punishA,choices=choices)

Kumar_fit <- stan(file = stanfile, data = Kumar_data, iter=chain_iter, chains = chain_n) #add working dir?
save(Kumar_fit, file=file.path(datadir,'Kumar_fit'))

traceplot(Kumar_fit,'lp__')

draws_Kumar<- extract(Kumar_fit)
summary_Kumar<- summary(Kumar_fit)

Kumar_loglike<- extract_log_lik(Kumar_fit, parameter_name = "loglik", merge_chains = TRUE)

loo(Kumar_loglike)

overall_summary<-summary_Kumar$summary #summary of all chains merged

#puts output in format for correlations
estimated_Kumar_output<- matrix(data=NA, nrow <- nsub_Kumar, ncol<- 8)
rownames<-paste('alpha[',1:nsub_Kumar,',1]',sep='')
estimated_Kumar_output[,1] <- overall_summary[c(rownames),'mean']
rownames<-paste('alpha[',1:nsub_Kumar,',1]',sep='')
estimated_Kumar_output[,2] <- overall_summary[c(rownames),'sd']
rownames<-paste('alpha[',1:nsub_Kumar,',2]',sep='')
estimated_Kumar_output[,3] <- overall_summary[c(rownames),'mean']
rownames<-paste('alpha[',1:nsub_Kumar,',2]',sep='')
estimated_Kumar_output[,4] <- overall_summary[c(rownames),'sd']
rownames<-paste('beta[',1:nsub_Kumar,',1]',sep='')
estimated_Kumar_output[,5] <- overall_summary[c(rownames),'mean']
rownames<-paste('beta[',1:nsub_Kumar,',1]',sep='')
estimated_Kumar_output[,6] <- overall_summary[c(rownames),'sd']
rownames<-paste('beta[',1:nsub_Kumar,',2]',sep='')
estimated_Kumar_output[,7] <- overall_summary[c(rownames),'mean']
rownames<-paste('beta[',1:nsub_Kumar,',2]',sep='')
estimated_Kumar_output[,8] <- overall_summary[c(rownames),'sd']

#for positive learning rate
cor.test(c(Kumar_awin_pat,Kumar_awin_con),estimated_Kumar_output[,1]) #correlation
lm(c(Kumar_awin_pat,Kumar_awin_con)~estimated_Kumar_output[,1]) #regression equation
plot(c(Kumar_awin_pat,Kumar_awin_con),estimated_Kumar_output[,1]) #plot

#for negative learning rate
cor.test(c(Kumar_aloss_pat,Kumar_aloss_con),estimated_Kumar_output[,3]) #correlation
lm(c(Kumar_aloss_pat,Kumar_aloss_con)~estimated_Kumar_output[,3]) #regression equation
plot(c(Kumar_aloss_pat,Kumar_aloss_con),estimated_Kumar_output[,3]) #plot

#for beta win
cor.test(c(Kumar_betawin_pat,Kumar_betawin_con),estimated_Kumar_output[,5]) #correlation
lm(c(Kumar_betawin_pat,Kumar_betawin_con)~estimated_Kumar_output[,5]) #regression equation
plot(c(Kumar_betawin_pat,Kumar_betawin_con),estimated_Kumar_output[,5]) #plot

#for beta loss
cor.test(c(Kumar_betaloss_pat,Kumar_betaloss_con),estimated_Kumar_output[,7]) #correlation
lm(c(Kumar_betaloss_pat,Kumar_betaloss_con)~estimated_Kumar_output[,7]) #regression equation
plot(c(Kumar_betaloss_pat,Kumar_betaloss_con),estimated_Kumar_output[,7]) #plot

```
